---
title: "R Code & Output"
author: "Huafeng Zhang"
date: "4/21/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Data Manpulation
```{r,message=FALSE,warning=FALSE}
setwd("~/Desktop/BIGbanks")
Banks16 <- read.csv("Banks16.csv", header = TRUE)
Assets <- read.csv("Assets16.csv",header = TRUE)
Banks16 <- left_join(Banks16, Assets, by = "cert")

# Add latitude & longitude for the U.S. map later 
library(zipcode)
# citation("zipcode")
library(dplyr)
# citation("dplyr")
data(zipcode)
z2 <- zipcode %>%
  mutate(zip = as.numeric(zip)) %>%
  select(zip, lat = latitude, long = longitude)
Banks16$zip<- as.numeric(Banks16$zip)
Banks16_geo_comp <- left_join(Banks16, z2, by="zip")
```

## Removed Banks with Missing Values 

```{r,message=FALSE,warning=FALSE}
# 85 NAs
sum(is.na(Banks16)) 
Banks16_draft <- Banks16_geo_comp[complete.cases(Banks16_geo_comp), ]

# Make the names of the five varables consitent with the terms I used in my paper
names(Banks16_draft)[12:16] <- c("NIM","ROA","NCOL", "CCLR", "TRBCR") 

# 48 banks was removed (5897-5849)
nrow(Banks16)-nrow(Banks16_draft)   
```

## PCP for the Data Set that is Used to Cluster (try to find potential skewed banks) 
```{r,message=FALSE,warning=FALSE}
library(lattice)
# citation("lattice")
parallelplot(Banks16_draft[ ,12:16],scale = list(x = list(rot = 45)),horizontal = FALSE)
```

## Reasoning for Excluding the Four Banks (P.S. the values of the assets are all in thousands (000's))
```{r,message=FALSE,warning=FALSE}
# Summary of all banks' assets
summary(Banks16_draft$asset_16) 

# The four banks' assets
TRBCR3 <- head(sort(Banks16_draft$TRBCR, decreasing = TRUE), 5)
ROA1 <-  head(sort(Banks16_draft$ROA, decreasing = TRUE), 5)
Bank4 <- filter(Banks16_draft, Banks16_draft$cert %in% c(27575,18923, 21761,33831))
Assets16_Bank4 <- Bank4$asset_16
Assets16_Bank4
```

## Removed the Skewed Banks 
```{r,message=FALSE,warning=FALSE}
Banks16_cont <- Banks16_draft[!Banks16_draft$cert %in% c(27575,18923, 21761,33831), ]
```

## Variable Summary for the Final Version Data Set 
```{r,message=FALSE,warning=FALSE}
library(knitr)
Banks16_var <- Banks16_cont[ ,-c(1:11, 17:19)]
colnames(Banks16_var) <- c("NIM","ROA","NCOL","CCLR", "TRBCR")
Variable <- c("NIM","ROA", "NCOL", "CCLR", "TRBCR" )
total<- rep(nrow(Banks16_cont),5) 
n_NAs <- as.integer(c(10,9,48,9,9)) # From summary(Banks16_geo_comp)
Size_withoutNAs< - as.integer(total-n_NAs)
Mean_withNAs <- round(c(3.680,1.0312 ,0.16777,11.677,26.310),digits = 3)
Mean_withoutNAs <- round(c(3.6983,0.9383,0.16767 ,11.274,18.840),digits = 3) # From summary(Banks16_var)
SD1 <- sqrt(var(Banks16_var$NIM))
SD2 <- sqrt(var(Banks16_var$ROA))
SD3 <- sqrt(var(Banks16_var$NCOL))
SD4 <- sqrt(var(Banks16_var$CCLR))
SD5 <- sqrt(var(Banks16_var$TRBCR))
SD_withoutNAs <- round(c(SD1,SD2,SD3,SD4,SD5),digits = 3)
var_sum <- data.frame(Variable,Size_withoutNAs,n_NAs,Mean_withNAs,Mean_withoutNAs,SD_withoutNAs)
knitr::kable(var_sum)
```

## Check Correlation between Variables
```{r,message=FALSE,warning=FALSE}
library(corrplot)
# citation("corrplot")
cor_banks <- cor(scale(Banks16_var[ ,1:5]))
knitr::kable(cor_banks, caption = "Correlation Between Variables in the Study")
corrplot(cor_banks,type = "lower") 
```

## Scale the Five Variables
```{r,message=FALSE,warning=FALSE}
scale_var <- scale(Banks16_var,center = TRUE, scale = TRUE)
```

## Find the Optimal NUmber of Clusters (the Elbow Method)
```{r,message=FALSE,warning=FALSE}
set.seed(66666)
c_df <- data.frame()
for (k in 1:20){
    c_w <- kmeans(scale_var,centers = k,iter.max = 100)
   c_df <- rbind(c_df,cbind(k,c_w$tot.withinss))
  }
names(c_df) <- c("Clutser","Total within SS")
plot(c_df,type = "b",xlab ="Number of clusters",ylab ="Within groups sum of squares",pch = 20,cex = 2) 
abline(v = 9,col = "blue")
```

## Run Clustering and Find the Cluster Summary
```{r,message=FALSE,warning=FALSE}
k9<- kmeans(scale_var,9)
Banks16_cont$Cluster <- k9$cluster
Cluster_means <- Banks16_cont %>%
                group_by(Cluster) %>%
                summarise(mean(NIM),mean(ROA),mean(NCOL),mean(CCLR),mean(TRBCR))
Cluster_means$Size <- c(k9$size)
knitr::kable(Cluster_means,caption = "Cluster Summary")
```

## PCP Sorted by Clusters
```{r,message=FALSE,warning=FALSE}
Banks16_var$Cluster <- k9$cluster
parallelplot(Banks16_var,col = Banks16_cont$Cluster,lty = Banks16_cont$Cluster,scale = list(x = list(rot = 45)),horizontal = FALSE)
```

## Banks Failed in 2017
```{r,message=FALSE,warning=FALSE}
# Data manipulation 
Failed <- read.csv("Failed.csv", header = TRUE)
Failed$Closing.Date <- as.character(Failed$Closing.Date)
Failed$Closing.Date <- as.character(Failed$Closing.Date)
substrRight <- function(x,n){
  substr(x, nchar(x)-n+1,nchar(x))
}
Failed$Year <- as.factor(substrRight(Failed$Closing.Date, 2))
colnames(Failed)[4] <- "cert"
fb <- select(Failed, cert, fYear = Year)
Banks16_cont_f <- left_join(Banks16_cont,fb, by ="cert")

# Find banks failed in 2017
Banks16_cont_f17 <- subset(Banks16_cont_f, fYear == "17")
View(Banks16_cont_f17[ ,-c(2:12)])

# Separate the origin clustering data set by using Banks failed in 2017 or not
Banks16_cont_Notf17 <- subset(Banks16_cont_f, fYear = NA)
```

## Visulize Banks in the U.S. Map
```{r,message=FALSE,warning=FALSE}
library(maps)
library(plotly)
# citation("maps")
# citation("plotly")
states <- map_data("state")
Banks16_cont_Notf17$Cluster <- as.factor(Banks16_cont_Notf17$Cluster)
Banks16_cont_f17$Cluster <- as.factor(Banks16_cont_f17$Cluster)
map1 <- ggplot() + 
  geom_polygon(data = states, aes(x = long, y = lat, fill = region, group = group), alpha = .1) +
  geom_point(data = Banks16_cont_Notf17, aes(x = long, y = lat, color = Cluster)) +
  geom_point(data = Banks16_cont_f17, aes(x = long, y = lat,size = 3)) +
  geom_text(label = Banks16_cont_f$Cluster,hjust = 0, vjust = 0) + scale_color_manual(values = c("blue", "grey", "yellow", "red", "lightgreen", "green", "orange","purple","lightblue")) +
  scale_size_continuous(name = "Failed in 2017") + 
  guides(fill = FALSE) + 
  theme_bw()
map1
# ggplotly(map1) 

## In order to see the banks that failed in 2017 in a more clearer way
Banks16_geo_cont_Notf17 <- filter(Banks16_cont_Notf17, long < -67 & long > -130  & lat>25 )
map2 <- ggplot() + 
  geom_polygon(data = states, aes(x=long, y=lat, fill=region, group=group), alpha=.1) +
  geom_point(data = Banks16_geo_cont_Notf17, aes(x = long, y = lat, color = Cluster)) +
  geom_point(data = Banks16_cont_f17, aes(x = long, y = lat,size = 3)) +
  geom_text(label= Banks16_cont_f$Cluster,hjust = 0, vjust = 0)+
  scale_color_manual(values=c("blue", "grey", "yellow", "red", "lightgreen", "green", "orange","purple","lightblue")) + scale_size_continuous(name = "Failed in 2017") + 
  guides(fill=FALSE) + 
  theme_bw()
map2
# ggplotly(map2)
```

